Covariate analysis of ROSMAP RNASeq

### 2016-04-19 code/Rmd/MetaAnalysis.Rmd
Minor modifications

# Utility function to download files from synapse
download.file <- function(id){ tmp = fread(synGet(id)@filePath, data.table=F, header=T);  }

@@

# Extract subset of differential expression data
DEXP$ROSMAP.PFC = DEXP$ROSMAP.PFC %>% 
  dplyr::filter(Comparison == "cogdx1-cogdx4") %>%
  dplyr::select(hgnc_symbol, P.Value)

DEXP$MSSB.FP = DEXP$MSSB %>% 
  dplyr::filter(Comparison == "BM_10.Control-BM_10.AD") %>%
  dplyr::select(hgnc_symbol, P.Value)

DEXP$MSSB.STG = DEXP$MSSB %>% 
  dplyr::filter(Comparison == "BM_22.Control-BM_22.AD") %>%
  dplyr::select(hgnc_symbol, P.Value)

DEXP$MSSB.PHG = DEXP$MSSB %>% 
  dplyr::filter(Comparison == "BM_36.Control-BM_36.AD") %>%
  dplyr::select(hgnc_symbol, P.Value)

@@

DEXP$MAYO.TC = DEXP$MAYO.TC %>% 
  dplyr::select(hgnc_symbol, P.Value)

DEXP$MAYO.CER.AR = DEXP$MAYO.CER.AR %>% 
  dplyr::select(hgnc_symbol, P.Value)

DEXP$MAYO.TC.AR = DEXP$MAYO.TC.AR %>% 
  dplyr::select(hgnc_symbol, P.Value)

GNames = DEXP[[1]]$hgnc_symbol
for (i in 2: length(DEXP)){
  GNames = intersect(GNames, DEXP[[i]]$hgnc_symbol)
}
GNames = GNames[!is.na(GNames)]
GNames = GNames[GNames != '']

@@

@@

writeLines('Genes that are differentially expressed at a FDR <= 0.05 in 6/8 datasets')
paste(rownames(Adj.P.Value)[which(rowSums(Adj.P.Value<=0.05)==6)], collapse = ',')
```

>>>>

# Utility function to download files from synapse
download.file <- function(id){ tmp = fread(synGet(as.character(id))@filePath, data.table=F, header=T);  }

@@

# Extract subset of differential expression data
DEXP$ROSMAP.PFC = DEXP$ROSMAP.PFC %>% 
  dplyr::filter(Comparison == "cogdx1-cogdx4") %>%
  dplyr::select(hgnc_symbol, P.Value, logFC)

DEXP$MSSB.FP = DEXP$MSSB %>% 
  dplyr::filter(Comparison == "BM_10.Control-BM_10.AD") %>%
  dplyr::select(hgnc_symbol, P.Value, logFC)

DEXP$MSSB.STG = DEXP$MSSB %>% 
  dplyr::filter(Comparison == "BM_22.Control-BM_22.AD") %>%
  dplyr::select(hgnc_symbol, P.Value, logFC)

DEXP$MSSB.PHG = DEXP$MSSB %>% 
  dplyr::filter(Comparison == "BM_36.Control-BM_36.AD") %>%
  dplyr::select(hgnc_symbol, P.Value, logFC)

@@

DEXP$MAYO.CER = DEXP$MAYO.CER %>% 
  dplyr::select(hgnc_symbol, P.Value, logFC)

DEXP$MAYO.TC = DEXP$MAYO.TC %>% 
  dplyr::select(hgnc_symbol, P.Value, logFC)

DEXP$MAYO.CER.AR = DEXP$MAYO.CER.AR %>% 
  dplyr::select(hgnc_symbol, P.Value, logFC)

DEXP$MAYO.TC.AR = DEXP$MAYO.TC.AR %>% 
  dplyr::select(hgnc_symbol, P.Value, logFC)
```

```{r plotDE}
tmp = lapply(DEXP, function(x){
  x$observed.P.val = p.adjust(x$P.Value, 'fdr')
  x = arrange(x, observed.P.val)
  x = mutate(x, expected.P.val = seq(0,1,1/(dim(x)[1]-1)))
  x$dexp = 0;
  x$dexp[x$observed.P.val <= 0.05 & x$logFC > 0] = 1;
  x$dexp[x$observed.P.val <= 0.05 & x$logFC < 0] = -1;
  dplyr::select(x, observed.P.val, expected.P.val, dexp)
})

p = list()
for (i in 1:length(tmp)){
  p[[i]] = ggplot(tmp[[i]], aes(x = -log10(expected.P.val), y = -log10(observed.P.val))) + geom_point(aes(color = factor(dexp))) 
  p[[i]] = p[[i]] + scale_colour_manual(values=c('green','grey','red')) + theme(legend.position="none")
  x = -log10(tmp[[i]]$expected.P.val)
  x[4]-x[1]/
  p[[i]] = p[[i]] + geom_abline(slope=1, intercept=0) + ggtitle(names(tmp)[i])
}
multiplot(plotlist = p, cols = 4)
```

```{r uniqGenes}
GNames = DEXP[[1]]$hgnc_symbol
for (i in 2: length(DEXP)){
  GNames = intersect(GNames, DEXP[[i]]$hgnc_symbol)
}
GNames = GNames[!is.na(GNames)]
GNames = GNames[GNames != '']

@@

writeLines('Genes that are differentially expressed at a FDR <= 0.05 in 8/8 datasets')
paste(rownames(Adj.P.Value)[which(rowSums(Adj.P.Value<=0.05)==8)], collapse = ',')

@@

writeLines('Genes that are differentially expressed at a FDR <= 0.05 in 6/6 RNASeq datasets')
paste(rownames(Adj.P.Value)[which(rowSums(Adj.P.Value[,-(4:5)]<=0.05)==6)], collapse = ',')
```

### 2016-04-19 code/Rmd/ROSMAP_RNASeq.Rmd
Minor modifications

## Load required libraries
library(synapseClient)
library(ggplot2)
library(reshape2)
library(data.table)
library(dplyr)
library(stringr)
library(limma)
library(knitr)
library(Biobase)
library(GEOquery)
library(RColorBrewer)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)
library(ggplot2)
library(WGCNA)

@@

@@

tree = hclust(as.dist(t(EXPR)))
plotDendroAndColors(tree, 
                    colors = labels2colors(COVARIATES.tmp), 
                    dendroLabels = FALSE, 
                    abHeight = 0.80, 
                    main = "Sample dendrogram",
                    groupLabels = colnames(COVARIATES.tmp))
```

@@

tree = hclust(as.dist(t(RESIDUAL.GENE_EXPRESSION[,-(1:2)])))
WGCNA::plotDendroAndColors(tree, 
                    colors = labels2colors(COVARIATES.tmp), 
                    dendroLabels = FALSE, 
                    abHeight = 0.80, 
                    main = "Sample dendrogram",
                    groupLabels = colnames(COVARIATES.tmp))
```

@@

DE = DE %>% plyr::ldply(.id = 'Comparison') %>%
  tidyr::separate(ensembl_gene_id, into=c("ensemble_id","position")) %>%
  dplyr::rename(ensembl_gene_id = ensemble_id) %>%
  left_join(Ensemble2HGNC)

>>>>

## Load required libraries
library(synapseClient)
library(ggplot2)
library(reshape2)
library(data.table)
library(dplyr)
library(stringr)
library(limma)
library(knitr)
library(Biobase)
library(GEOquery)
library(RColorBrewer)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)
library(ggplot2)

@@

plotdata <- left_join(plotdata, METADATA) %>%
  dplyr::mutate(Batch = factor(Batch),
                age_death = as.numeric(age_death),
                cogdx = factor(cogdx))

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Batch, shape=cogdx, size=age_death))
p <- p + theme_bw() + theme(legend.position="top") 
# p <- p + geom_text(aes(label= Sampleid), size=4, hjust=1)
p <- p + xlab(paste0('PC1(',round((PC$sdev^2/sum(PC$sdev^2))[1]*100),'%)'))
p <- p + ylab(paste0('PC2(',round((PC$sdev^2/sum(PC$sdev^2))[2]*100),'%)'))
p

@@

tree = hclust(as.dist(t(EXPR)))
plotDendroAndColors(tree, 
                    colors = labels2colors(COVARIATES.tmp), 
                    dendroLabels = FALSE, 
                    abHeight = 5, 
                    main = "Sample dendrogram",
                    groupLabels = colnames(COVARIATES.tmp))
```

@@

tree = hclust(as.dist(t(RESIDUAL.GENE_EXPRESSION[,-(1:2)])))
plotDendroAndColors(tree, 
                    colors = labels2colors(COVARIATES.tmp), 
                    dendroLabels = FALSE, 
                    abHeight = 0.80, 
                    main = "Sample dendrogram",
                    groupLabels = colnames(COVARIATES.tmp))
```

@@

DE = DE %>% plyr::ldply(.id = 'Comparison')
